---
title: "Spotify_songs_analysis"
author: "Horgonyi Anna"
date: "`r Sys.Date()`"
output: html_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This project analyzes the Spotify Songs dataset from the TidyTuesday project. The main objectives are to perform exploratory data analysis, build and compare statistical models, and provide human-readable and well-documented results.

## Data Import and Preparation

```{r load-packages}
# Set the CRAN mirror
chooseCRANmirror(graphics = FALSE, ind = 1) # This chooses the first CRAN mirror

# Install necessary packages
install.packages("janitor")
install.packages("caret")
install.packages("broom")
install.packages("knitr")

# Load the libraries
library(tidyverse)
library(janitor)
library(ggplot2)
library(caret)
library(broom)
library(knitr)

```

```{r load-data}
# import data from GitHub
url <- "https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2020/2020-01-21/spotify_songs.csv"
spotify_songs <- read_csv(url) %>% 
  clean_names()

# quick overview of the dataset
head(spotify_songs)
summary(spotify_songs)
```

## Exploratory Data Analysis (EDA)

### Distribution of Key Variables

```{r eda-plots}

# relationship between acousticness and energy
spotify_songs %>% 
  ggplot(aes(x = acousticness, y = energy)) + 
  geom_point(alpha = 0.3) + 
  geom_smooth(method = "lm", color = "red") + 
  labs(title = "űacousticness vs. Energy", x = "Acousticness", y = "Energy")

# relationship between speechiness and liveness
spotify_songs %>% 
  ggplot(aes(x = speechiness, y = danceability)) + 
  geom_point(alpha = 0.3) + 
  geom_smooth(method = "lm", color = "red") + 
  labs(title = "Speechiness vs. Liveness", x = "Speechiness", y = "Liveness")
```

## Statistical Models

### Model 1: predicting danceability using energy and acousticness
#### Hypothesis: Both energy and acousticness have a significant positive relationship with danceability. 

```{r model1}
# create model 1
model1 <- lm(danceability ~ energy + acousticness, data = spotify_songs)

# display summary of model 1
summary(model1)

# check residuals of model 1
par(mfrow = c(2, 2))
plot(model1)

# perform normality check using shapiro-wilk test on residuals
shapiro_test <- shapiro.test(sample(resid(model1), 5000))
shapiro_test


```

### Model 2: predicting liveness using valence, tempo, and speechiness
#### Hypothesis: Valence, tempo, and speechiness significantly predict the liveness of songs.

```{r model2}
# create model 2
model2 <- lm(liveness ~ valence + tempo + speechiness, data = spotify_songs)

# display summary of model 2
summary(model2)

# check residuals of model 2
par(mfrow = c(2, 2))
plot(model2)

# perform normality check using shapiro-wilk test on residuals
shapiro_test <- shapiro.test(sample(resid(model2), 5000))
shapiro_test

# plots show a "funnel-shape" so checking for homoscedasticity with the breusch-pagan test

library(lmtest)
bptest(model1)

# homoskedasticity violated, do corrections

# load the necessary package
install.packages("sandwich")
library(sandwich)

# calculate robust standard errors
robust_se <- sqrt(diag(vcovHC(model1, type = "HC1")))

# display the robust standard errors
robust_se

```

## Model Evaluation

```{r model-evaluation}
# AIC and BIC for Model 1
aic_model1 <- AIC(model1)
bic_model1 <- BIC(model1)

# AIC and BIC for Model 2
aic_model2 <- AIC(model2)
bic_model2 <- BIC(model2)

# Output the results
aic_model1
bic_model1
aic_model2
bic_model2
```
### Conclusions
##### Model 1 has lower AIC and BIC values compared to model 2, so model 2 is a better fit for the data than model 1. The hypothesis that both energy and acousticness have a positive relationship with danceability is not supported by the results. Both predictors show a negative relationship with danceability. The statistical significance of these predictors (with very small p-values) indicates that they are indeed related to danceability, but their effect is negative, which contradicts the initial expectation. The R-squared value is low, suggesting that the model does not explain much of the variability in danceability. This could indicate that other factors, not included in this model, may be more important in predicting danceability.




